Time Series Analysis of Chaotic Dynamical Systems with Recurrent Neural Networks

### 2019-04-17 doc/paper/paper.tex

>>>>

The double pendulum's initial condition was

\begin{equation}
(\theta_1, \theta_2, \omega_1, \omega_2) 
= (4.66648088, 4.29820501,
8.01585663 \times 10^{-4}, 3.71175422 \times 10^{-3}).
\end{equation}
In this case, we
need $\theta_1$ large enough that the system is chaotic, but the rest of
the parameters were chosen randomly (with rotational velocities close to
zero). The resulting Lyapunov Exponent is $0.60074$. 
Like the Lorenz, the differences between time points
were used to train the ESN since the ESN was showing poor results
otherwise. As the pendulum kept looping the center, $\theta_1$ and 
$\theta_2$ kept growing; this shouldn't be an issue, but using
the differences keeps the system variables in a closer range.

Figure \ref{fig:doub_pend_diff} shows a plot of each
of the dataset's dimensions. It seems like the dataset's activity is
fairly sparse. There is only gradual but seemingly random variations until
a few spots around $t=10, 30, 60, \text{and} 75$.
The results for the ESN weren't very strong, but it was able to predict
almost two Lyapunov Times before diverging.

Because there are portions of the dataset which are calm, the ESN then has
to pickup on signals which lead to the spikes in movement.
We see the testing graph that while it predicts the testing set well,
it eventually tries to predict spikes but does so at the wrong time,
and the spike pattern never ends, has too large of magnitude, and is very 
rapid. Overall the model performs poorly, and it's reflected in the RMSE.
The likely reason for this is that the ESN can't learn the complex
patterns from such a small and sparse dataset.

\begin{figure}[H]
    \centering
    \includegraphics[width=.5\linewidth]{doc/paper/images/doub_pend/full_differential.png}
    \caption{The entire difference dataset}
    \label{fig:doub_pend_diff}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=.5\linewidth]{doc/paper/images/doub_pend/test_data.png}
    \caption{The testing dataset partition}
    \label{fig:doub_pend_testing}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_0_param_77_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_0_param_225_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_1_param_63_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_1_param_230_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_2_param_305_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_2_param_458_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_3_param_52_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_3_param_151_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/doub_pend/rank_4_param_127_fit.png}
        \caption{Delayed spike prediction}
    \end{subfigure}
    ~
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{doc/paper/images/lorenz/rank_4_param_300_rmse.png}
        \caption{RMSE Over Time}
    \end{subfigure}
    \caption{Poor forecast of double pendulum}
\end{figure}
